---
title: "Homework 1 Solution"
output: github_document
---

```{r setup, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Context

This assignment reinforces ideas in the [building blocks](topic_building_blocks.html) topic. 

## Due date and submission

Due: September 29 at 10:00pm. 

Please submit (via courseworks) the web address of the GitHub repo containing your work for this assignment; git commits after the due date will cause the assignment to be considered late.

R Markdown documents included as part of your solutions must not install packages, and should only load the packages necessary for your submission to knit. 

## Points

| Problem         | Points    |
|:--------------- |:--------- |
| Problem 0.1     | 25        |
| Problem 0.2     | 25        |
| Problem 1       | 25        |
| Problem 2       | 25        |
| Optional survey | No points |


## Problem 0.1

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.


## Problem 0.2

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 


## Problem 1

## Problem 1

### Create data frame

Below we create a tibble consisting of:

* a random sample of size 10 from a standard Normal distribution
* a logical vector indicating whether elements of the sample are greater than 0
* a character vector of length 10
* a factor vector of length 10, with 3 different factor "levels"

```{r create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
  )
```

### Take the mean of each variable

When we try to take the mean of each variable:

* We are successful with the variable `samp` because it is a numeric value. 
* We are successful with the variable `samp_gt_0` because R can interpret the logical vector as a vector of 0's (FALSE's) and 1's (TRUE's). The resulting mean is the proportion of TRUE values.
* We get warnings for the variables `char_vec` and `factor_vec` because neither are numeric nor logical, and can't be interpreted as numeric. The resulting output is `NA`.

```{r mean_pull_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

### Convert variables to numeric

Next we apply the `as.numeric` function to the logical, character, and factor variables.

* We successfully convert `samp_gt_0` to numeric, taking values of 0 (FALSE) and 1 (TRUE).
* We cannot convert `char_vec` to numeric. The result is a warning and a vector of NA's.
* We successfully convert `factor_vec` to numeric, taking values of 2 for `"low"`, 3 for `"moderate"`, and 3 for `"high"`. Numeric values are assigned to the factor levels, which are alphabetical unless otherwise specified.

We show the code but not the output below by specifying code chunk options `results="hide"`, `warning=FALSE`, and `message=FALSE`.

```{r try_numeric, results = "hide", warning=FALSE}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```



## Problem 2

### Load dataset

The package is installed, and does not need to be reinstalled as part of this document. The code below loads the data.

```{r}
data("penguins", package = "palmerpenguins")
```

After loading, some data inspections are encouraged (e.g. using `view` or reading a help file about the data). These are also not included in this document.

### Data description
 
The dataset contains observations on `r nrow(penguins)` individual penguins from three species living on three islands. The data include `r ncol(penguins)` variables, and are primarily physiological -- measurements of bill, flippers, body mass, and sex -- in addition to species, island, and year.
 
The mean flipper length (rounded to the nearest mm) of penguins with nonmissing values in this dataset is `r round(mean(pull(penguins, flipper_length_mm), na.rm = TRUE))` mm.
 
Below is a scatterplot comparing flipper length to bill length, color coded by species. 

```{r scatterplot}
ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) + 
  geom_point()

ggsave("p8105_hw1_ajg2202.pdf")
```

### Optional survey

Optional survey is optional. 

